---
title: 'STA 602 Homework and Lab Template'
author: "Student"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
---

```{r setup, message=F, warning=F, echo=F}
library(tidyverse)
require(magrittr)
require(plyr)
ggplot2::theme_set(ggplot2::theme_bw())
knitr::opts_chunk$set(fig.align = 'center')
```

Here's a list of resources for learning rmarkdown:

1. [R markdown introduction](https://rmarkdown.rstudio.com/lesson-1.html)
1. [R for Data Science](https://r4ds.had.co.nz/)
1. [Introduction to R markdown](https://rmarkdown.rstudio.com/articles_intro.html)
1. [Another introduction to R markdown](https://rpubs.com/andrew-cho/303981)
1. [R markdown cheatshett](https://rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf)
1. [Data visualization with ggplot2 cheat sheet](https://rstudio.com/wp-content/uploads/2016/11/ggplot2-cheatsheet-2.1.pdf)
1. [Learn LaTex in 30 minutes](https://www.overleaf.com/learn/latex/Learn_LaTeX_in_30_minutes)


***

## Exercise 1

Brief statement of the problem (optional)

### Part (a)

Then, some math:

$$
\begin{aligned}
X \sim N(\mu,1) \implies p_X(x) = \frac{1}{\sqrt{2\pi}}e^{-\frac{1}{2}(x - \mu)^2}
\end{aligned}
$$

### Part (b)

Finally, some code:

```{r, warning = F}
x <- rnorm(1000, 0, 1)
x %>% data.frame() %>% ggplot2::ggplot() + geom_histogram(aes(x = x))
  
```

## Exercise 2

Repeat...

### Part (a)

### Part (b)

***

