---
title: "Lab 1: R review"
author: "Ryan Tang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  tufte::tufte_html:
    tufte_variant: "envisioned"
    highlight: pygments
    css: lab.css
    toc: true
link-citations: yes
---
```{r include=FALSE}
library(tidyverse)
```


### Exercise 1

1. Create a code chunk and set the header parameter to TRUE and print out the top rows of the table with `head()` as above.

```{r include=TRUE}
fpath <- "http://www2.stat.duke.edu/~pdh10/FCBS/Exercises/azdiabetes.dat"

data <- read.table(file = url(fpath), header = TRUE)
head(data)
```


### Exercise 2

2. Generate a sequence of $100$ equispaced real numbers from $0$ to $1$ and store it in a variable called `seq2`.
```{r include=TRUE}
seq2 <- seq(1, 100, by = 1)
seq2
```


### Exercise 3

3. Sort the entries in `seq3` from greatest to least.

```{r include=TRUE}
seq3 <- seq(from = -3, to = 3, by = .5)
sort(seq3, decreasing = TRUE)
```


### Exercise 4

4. Find the variance of each row of `mat5`

```{r include=TRUE}
mat5 <- matrix(seq(1, 100, 1), nrow = 4, ncol = 25, byrow = T)

mat5 %>%
  apply(MARGIN = 1, FUN = var)
```


### Exercise 5

5. Generate $500$ samples from a Beta distribution with shape parameter $[a, b] = [0.5, 0.5]$ and store the samples in a variable called `W`

```{r include=TRUE}
W <- rbeta(500, shape1=0.5, shape2=0.5)
W
```


### Exercise 6

6. Browse online resources (some below), or use code from above to make a few plots of your own.

```{r include=TRUE}
samples <- rnorm(1000000, mean = 0, sd = 1)

hist(samples, main = "Standard Normal Distribution")

plot(density(samples))

samples %>%
  data.frame(x = .) %>%
  ggplot2::ggplot() +
  geom_histogram(aes(x = x, y = ..density..), 
                 fill = "#756bb1", colour = "white", 
                 alpha = 0.5, bins = 30) +
  geom_density(aes(x = x), colour = "#756bb1") +
  labs(x = "x", y = "Density", title = "ggplot density / histogram")
```