---
title: 'STA 602 Lab 5'
author: "Ryan Tang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
---

```{r setup, message=F, warning=F, echo=F}
require(tidyverse)
require(rstanarm)
require(magrittr)
require(rstan)
require(plyr)
require(bayesplot)
require(loo)
require(readxl)
require(ggplot2)

setwd(".")
ggplot2::theme_set(ggplot2::theme_bw())
knitr::opts_chunk$set(fig.align = "center")
```

***
## Exercise 1
+ Exp(1) cannot be used as a candidate distribution to draw samples from N(0,1) on the interval
(-3, 3), because it has no support on the negative plane.
+ Between the interval of (-2, 2), N(0, 1) can be used to draw standard Cauchy samples because
we have a finite ceiling constant. The standard Cauchy has the highest density, $\frac{1}{\pi}$
at the point of x=0; same for the standard normal with a density of $\frac{1}{\sqrt{2\pi}}$ at
the same location. Hence, we can always find a scaling constant, $M$, for the rejection sampling.
+ Yes, knowing the kernel is enough for rejection sampling. We can draw the samples and then
normalize it using the finite samples. It is a Yes for a 100-variate distribution.
However, the efficiency is not ideal.

## Exercise 2
Given just the kernel is the possible $M'$ are anything that is $\geq 1$ for a uniform candidate
distribution. The range is certainly different if apply to the true density function.

## Exercise 3
```{r}
# target density function
f <- function(x) {
  sin(pi * x)^2
}

set.seed(23523)
nsim <- 10000
M <- 1
accepted_count <- 0
total_count <- 0
res <- numeric()

while (accepted_count < nsim){
  x <- runif(1)
  r <- f(x) / (M * dunif(x))
  if (runif(1) <= r) {
    res <- append(res, x)
    accepted_count <- accepted_count + 1
  }

  total_count <- total_count + 1
}

theta <- seq(0, 1, length.out = 200)
g_theta <- sin(pi * theta)^2
ggplot() +
  geom_density(data=data.frame(xs = res), aes(xs)) +
  geom_line(data=data.frame(theta=theta, g=g_theta), aes(theta, g), color="red")
```

## Exercise 4
+ Because $sin^2(\pi x)$ is no the real density, just a kernel, of course the samples we obtained from the rejection sampler will differ.
But the overall shape is simliar.
+ At $M=1$, I had about 50% acceptance rate. A total counts of 20,026 for 10,000 accepted samples.
+ The higher the $M'$ is, the lower the acceptance rate. On the contrary, if $M'$ is too small, the candidate q will not cover the entire support
of the target distribution, we will not obtain good samples.

```{r echo=F, include=T}
# target density function
f <- function(x) {
  sin(pi * x)^2
}

set.seed(23523)
nsim <- 10000
M <- 3
accepted_count <- 0
total_count <- 0
res <- numeric()

while (accepted_count < nsim){
  x <- runif(1)
  r <- f(x) / (M * dunif(x))
  if (runif(1) <= r) {
    res <- append(res, x)
    accepted_count <- accepted_count + 1
  }

  total_count <- total_count + 1
}

theta <- seq(0, 1, length.out = 200)
g_theta <- sin(pi * theta)^2
ggplot() +
  geom_density(data=data.frame(xs = res), aes(xs)) +
  geom_line(data=data.frame(theta=theta, g=g_theta), aes(theta, g), color="red") +
  ggtitle("M = 3")
```

```{r echo=F, include=T}
# target density function
f <- function(x) {
  sin(pi * x)^2
}

set.seed(23523)
nsim <- 10000
M <- 0.25
accepted_count <- 0
total_count <- 0
res <- numeric()

while (accepted_count < nsim){
  x <- runif(1)
  r <- f(x) / (M * dunif(x))
  if (runif(1) <= r) {
    res <- append(res, x)
    accepted_count <- accepted_count + 1
  }

  total_count <- total_count + 1
}

theta <- seq(0, 1, length.out = 200)
g_theta <- sin(pi * theta)^2
ggplot() +
  geom_density(data=data.frame(xs = res), aes(xs)) +
  geom_line(data=data.frame(theta=theta, g=g_theta), aes(theta, g), color="red") +
  ggtitle("M = 0.25")
```


## Exercise 5
Beta(2, 2) is a nice choice compare to the uniform candidate. Using this, the acceptance rate
increased from 50% to around 70%. 

```{r echo=F, include=T}
# target density function
f <- function(x) {
  sin(pi * x)^2
}

set.seed(23523)
nsim <- 10000
M <- 1/1.5
accepted_count <- 0
total_count <- 0
res <- numeric()

while (accepted_count < nsim){
  x <- rbeta(1, shape1=2, shape2=2)
  r <- f(x) / (M * dbeta(x, shape1=2, shape2=2))
  if (runif(1) <= r) {
    res <- append(res, x)
    accepted_count <- accepted_count + 1
  }

  total_count <- total_count + 1
}

theta <- seq(0, 1, length.out = 200)
g_theta <- sin(pi * theta)^2
ggplot() +
  geom_density(data=data.frame(xs = res), aes(xs)) +
  geom_line(data=data.frame(theta=theta, g=g_theta), aes(theta, g), color="red") +
  ggtitle("q = Beta(2, 2) | M = 1/1.5")
```
***