---
title: 'STA 602 Lab 7'
author: "Ryan Tang"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
---

```{r setup, message=F, warning=F, echo=F}
library(tidyverse)
library(rstanarm)
library(magrittr)
library(rstan)
library(plyr)
library(bayesplot)
library(loo)
library(readxl)
library(ggplot2)
library(ggrepel)
library(cowplot)
library(truncnorm)

setwd(".")
ggplot2::theme_set(ggplot2::theme_bw())
knitr::opts_chunk$set(fig.align = "center")
```

***
## Exercise 1
Here we sample from the derived full conditional equations.
```{r echo=F}
n <- 1000
true_Rsquared <- 5
true_sigma <- 1.25
#
u <- runif(n, 0, true_Rsquared)  # de-noised squared distance of each point to (0,0)
r <- u + rnorm(n, sd = true_sigma)  # add normal noise
theta <- runif(n, 0, 2*pi)
#
ggplot2::ggplot() +
  geom_point(data = data.frame(x = sign(r)*sqrt(abs(r))*cos(theta), y = sign(r)*sqrt(abs(r))*sin(theta)),
             aes(x = x, y = y), shape = 1) +
  geom_path(data = data.frame(R = true_Rsquared) %>%
                              plyr::ddply(.(R), function(d){
                                data.frame(x = sqrt(d$R)*cos(seq(0, 2*pi, length.out = 100)),
                                            y = sqrt(d$R)*sin(seq(0, 2*pi, length.out = 100)))
                              }),
                       aes(x = x, y = y), alpha = 1, colour = "red") +
  coord_fixed()
```

```{r echo=T}
set.seed(23498324)

# hyper-parameters
m <- 3
k <- 1
alpha <- 5/2
beta <- 5/2

# generate random sample from pareto(m,k)
rpareto <- function(m, k, trunc = NULL){
  p <- m*(1 - runif(1))^(-1/k)
  if(!is.null(trunc)){
    while(p > trunc){
      p <- m*(1 - runif(1))^(-1/k)
    }
  }
  return(p)
}

#
uni_pareto_gibbs <- function(S, r, m, k, alpha, beta, burn_in = min(1000, S / 2), thin = 5){
  # Reparametrize X matrix to squared radius values
  Rsq <- r  # squared distance
  n <- length(Rsq)  # sample size
  R2 <- rep(1, S)  # to save the S gibbs samples for R^2, also intialized it
  U <- matrix(0, nrow = S, ncol = n)  # to save the S gibbs samples for u_1, ..., u_n
  U[1, ] <- runif(n, 0, R2)  # initialize the first sample for u_1, ..., u_n
  sigma <- rep(1, S)  # to save the S gibbs samples for sigma (not sigma2!)

  # initialize
  U_curr <- U[1, ]  # save the updated sample for u_1, ..., u_n
  R2_curr <- R2[1]  # save the updated sample for R^2
  sigma_curr <- sigma[1]  # save the updated sample for sigma (not sigma2!)
  for(s in 1:S){
    # Sample from full conditional of the inner radius
    R2_curr <- rpareto(max(c(U_curr, m)), k + n)
    R2[s] <- R2_curr
    # Sample from full conditional of U values
    U_curr <- truncnorm::rtruncnorm(n, a = 0, b = R2_curr, mean = Rsq, sd = sigma_curr)
    U[s, ] <- U_curr
    # Sample from full conditional of sigma
    sigma_curr <-  sqrt(1 / rgamma(1, shape=n/2+alpha, rate=beta + 0.5*sum((r-U_curr)^2)))
    sigma[s] <- sigma_curr
  }
  return(list(R = R2[seq(burn_in, S, by = thin)], 
              U = U[seq(burn_in, S, by = thin), ], 
              sigma = sigma[seq(burn_in, S, by = thin)]))
}

gibbs_samps <- uni_pareto_gibbs(S = 100000, r, m, k, alpha, beta, burn_in=2000)
```


# Exercise 2
Here we plot the inferred $R^2$ posterior marginal against the true $R^2$. It looks like we
are doing an excellent job of estimating the mean and the shape. 
```{r echo=T}
ggplot() +
  geom_density(data = data.frame(R = gibbs_samps$R), aes(x=R)) +
  geom_vline(xintercept = true_Rsquared, color='#da1a1a', linetype="dotted")
```

# Exercise 3
Here we plot the inferred $\sigma^2$ posterior marginal against the true $sigma^2$.
On average, we are underestimating the noise term here.
```{r echo=T}
ggplot() +
  geom_density(data = data.frame(sigma = gibbs_samps$sigma^2), aes(x=sigma)) +
  geom_vline(xintercept = true_sigma^2, color='#da1a1a', linetype="dotted")
```

# Exercise 4
Furthermore, here we do a 2D contour plot between $(R^2, \sigma^2)$ and compare the distribution to
the true parameter. Matching with the previous comments that we are underestimating the variances.
```{r echo=T}
ggplot() +
  geom_density2d_filled(data = data.frame(R = gibbs_samps$R, sigma = gibbs_samps$sigma^2), aes(sigma, R)) +
  geom_vline(xintercept = true_sigma^2, color='#da1a1a', linetype="dotted") +
  geom_hline(yintercept = true_Rsquared, color='#da1a1a', linetype="dotted")
```
***
